FIXME

### 2013-08-30 writeups/proposal/proposal.tex
foo
Former-commit-id: 6f64ec51c5a212be9d94e8dcca38b420656a4bf8 [formerly 01f3ffd4837d4cba3d282174f29d3182618f215b]
Former-commit-id: 377d4ec6086f9fef22b6fa8b5c54324a1fae265c

TODO: Summarize \cite{Plate1995}

@@

>>>>

``holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., chunk) as a vector.
The representation of the concept is distributed across all elements of the vector.
This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and saying that the distribution of co-occurrence counts across the rows in the column is the representation of that concept.
However, vector-based memory systems represent information in a much more compact way than a full word co-occurrence matrix.

For vector-based systems, each word is represented by an environment vector $e_{j}$ that is nearly orthogonal to all other words' environment vectors.
The number of dimensions for these vectors is much less than the number of rows in a full word co-occurrence matrix, which is how vector-based systems represent information in a much more compact space.
Paired with each environment vector ($e_{i}$) is a memory vector ($m_{i}$) that contains the summed representation of all other environment vectors that have co-occured with that $e_{i}$ environment vector.
These memory vectors can contain environment vectors from bag-of-words co-occurrence information ($c_{i}$) and word order information ($o_{i}$), which can be combined into a single representation ($m_{i}$).
Over time, memory vectors accumulate a distributed representation of the most common environment vectors that co-occurred with them.

\subsubsection{Retrieval Process}

Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
Decoding takes the memory vector for context and decodes it back into an environment vector.
The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.
For example, assume that a user has written the word ``zend'' in a post.
``zend'' has co-occurred with \emph{PHP}, so the unordered memory vector for ``zend'' ($c_{zend}$) contains the unordered environment vector for \emph{PHP}.
Since the environment vector for \emph{PHP} was added to the memory vector for ``zend'' without any encoding, no decoding is necessary.
So the cosine between the environment vector \emph{PHP} and decoded memory vector for ``zend'' ($c_{zend}$) will be high, and most likely \emph{PHP} will be retrieved through decoding.

Resonance is the opposite retrieval process, where a memory vector is created from the context, and then that context memory vector is compared to all memory vectors.
Fo the \emph{PHP} and ``zend'' example, the context memory vector would consist of the unordered environment vector for ``zend''.
Since the memory vector for \emph{PHP} contains the ``zend'' environment vector, it is likely that the correlation between these two memory vectors will be high.

Usually both resonance and decoding will return very similar rank orderings of most likely chunks given context.
\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
However, decoding does require an additional decoding step that resonance does not require, and depending on the decoding operation, this step may be computatinoally expensive.

@@

One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
Representing word order is an example where the endoding and decoding operations are applied.
As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
Simply counting co-occurrences of the two immediately-preceding words to hashtags does not include the order information of the words in the representation.
With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occured just prior to these 1-back words.
Order information is lost if a simply co-occurrence counting technique is used.

With vector-based systems, order informatino is included by creating an environment vector ($e_{bind(i,\Phi)}$) that is a function of environment vectors for both the word ($e_{i}$) and location ($\Phi$).
This function has the useful property that the resulting environment vector and the location vector can be inverted to return the original word's environment vector.
Convolution and inverse convolution (i.e,. circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007} to implement these operations.
This allows for retrieval requests such as ``playing \#$\Phi$'', ``just read \#$\Phi$'', and ``vector-based $\Phi$ systems''.

For example, suppose that a retrieval request is made for ``martin $\Phi$'' given that the model was trained on the TASA corpus.
Within this corpus, the phrase ``martin luther'' often occured.
When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated with the convolving luther and a 1-back location vector ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin and a 1-forward location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.

Using decoding, the word-order memory vector for martin is deconvolved to the right ($o_{martin} \circledast \Phi$) which returns the environment vector for luther ($e_{luther}$).
Using resonance, a memory vector for ``martin $\Phi$'' is constructed ($e_{martin} * \Phi$), which returns the word-order memory vector for luther ($o_{luther}$).
Either way, luther is returned as the most likely term to match the phrase ``martin $\Phi$''.

New vectors generated by circular convolution are uncorrelated with all other environment vectors and have the same length.
This allows the word order information and unordered context information for a chunk to be aggregated and represented in the same memory vector.
Quite simply, the different types of memory vectors can be summed to create a single memory vector representation for each chunk (i.e., $m_{j} = c_{j} + o_{j}$).
